Quotient Normalized Maximum Likelihood Criterion for Learning Bayesian Network Structures

### 2018-03-26 poster/qnmlPoster.tex
polishing

\begin{block}{Background: Structure Learning of Bayesian Networks}
\begin{columns}[T]
   \begin{column}{0.3\textwidth} % first column {{{
     \heading{Bayesian Networks}
     \vspace*{12pt}
     \begin{itemize}
     \setlength\itemsep{1em}
     \item Provide a compact way to represent a joint distribution over a random vector $X = (X_1, \ldots , X_d)$. 
     \item Consist of:
     \begin{itemize}
     \item[1.] A Directed acyclic graph $G$ which encodes the dependencies between the components of $X$.
     \item[2.] Parameters $\theta = (\theta_1,\ldots,\theta_d)$, where $\theta_i$ denotes the parameters of the conditional distribution of $X_i$ given its parents $G_i$. 
     \end{itemize}
     \item Decomposition: $P(X \mid G, \theta) = \prod_{i=1}^dP(X_i\mid X_{G_i},\theta_i)$
     \end{itemize}     
   \end{column}

@@

	\item However, for \textbf{a single categorical data vector}, regret and one dimensional $P^1_{NML}$-code can be computed.
	\item \textbf{Trick}. Recall factorization w.r.t. graph $G$:
	$$
	P(D\mid G) = \prod_{i = 1}^{d}P(D_i \mid D_{G_i}) = \prod_{i = 1}^{d}\frac{P(D_{i,G_i})}{P(D_{G_i})}
	$$
	\item Assume there are \textbf{no independencies} among $G_i$. Now, $D_{i,G_i} = $
	\vspace*{12pt}
	\end{itemize}
	\begin{minipage}{0.45\textwidth}
	\begin{align*}
	\begin{blockarray}{c|ccc}
	D_i\phantom{s} & \phantom{..}  & D_{G_i} & \phantom{..}\\
	\begin{block}{(c|ccc)}
  	1 & 1 & 1 & 1  \\
  	0 & 1 & 0 & 0 \\
  	1 & 1 &0 & 0  \\
  	1 & 1 & 1 & 1  \\
	\end{block}
	\end{blockarray} 
	\ \to \ 
	\begin{blockarray}{c}
	\phantom{...}\\
	\begin{block}{(c)}
  	0   \\
  	1  \\
  	2  \\
  	0   \\
	\end{block}
	\end{blockarray}
 	\end{align*}
	\end{minipage}\quad
	\begin{minipage}{0.45\textwidth}
	\shadowbox{
	\parbox{\textwidth}{
	\vspace*{12pt}
	\begin{itemize}
	\item We can treat $D_{i,G_i}$ and $D_{G_i}$ as values of single variables!
	\item $P(D_{i,G_i})$ and $P(D_{i,G_i})$ can be modeled using $P_{NML}^1$-code.
	\end{itemize}}}  
	\end{minipage}

@@

\vskip-1.0ex
\begin{columns}[T]
  \begin{column}{0.63\paperwidth}
\begin{block}{References} %{{{
%\scriptsize
\tiny
{
\bibliographystyle{apalike}
\bibliography{../cosco}
}
\end{block} %}}}
  \end{column}
  \begin{column}{0.33\paperwidth}
\begin{block}{Acknowledgements} %{{{
%\scriptsize
\tiny
{
J.L., E.J. and T.R. were supported by the Academy of Finland (COIN CoE and Project \textsc{TensorML}). J.L. was supported by the DoCS doctoral programme of the University of Helsinki.
}
\end{block} %}}}
  \end{column}
\end{columns}
\end{frame}
\end{document}

>>>>

\begin{block}{Background: Structure Learning of Bayesian Networks}
\begin{columns}[T]
   \begin{column}{0.3\textwidth} % first column {{{
     \heading{Bayesian Networks}
     \vspace*{12pt}
     \begin{itemize}
     \setlength\itemsep{1em}
     \item Provide a compact way to represent a joint distribution over a random vector $X = (X_1, \ldots , X_d)$. 
     \item Consist of:
     \begin{itemize}
     \item[1.] A Directed acyclic graph $G$ which encodes the dependencies between the components of $X$.
     \item[2.] Parameters $\theta = (\theta_1,\ldots,\theta_d)$, where $\theta_i$ denotes the parameters of the conditional distribution of $X_i$ given its parents $X_{G_i}$. 
     \end{itemize}
     \item Decomposition: $P(X \mid G, \theta) = \prod_{i=1}^dP(X_i\mid X_{G_i},\theta_i)$
     \end{itemize}     
   \end{column}

@@

	\item However, for \textbf{a single categorical data vector}, regret and one dimensional $P^1_{NML}$-code can be computed.
	\item \textbf{Trick}. Recall factorization w.r.t. graph $G$:
	$$
	P(D\mid G) = \prod_{i = 1}^{d}P(D_i \mid D_{G_i}) = \prod_{i = 1}^{d}\frac{P(D_{i,G_i})}{P(D_{G_i})}
	$$
	\item Assume there are \textbf{no independencies} among $X_{G_i}$. Now, $D_{i,G_i} = $
	\vspace*{12pt}
	\end{itemize}
	\begin{minipage}{0.45\textwidth}
	\begin{align*}
	\begin{blockarray}{c|ccc}
	D_i\phantom{s} & \phantom{..}  & D_{G_i} & \phantom{..}\\
	\begin{block}{(c|ccc)}
  	1 & 1 & 1 & 1  \\
  	0 & 1 & 0 & 0 \\
  	1 & 1 &0 & 0  \\
  	1 & 1 & 1 & 1  \\
	\end{block}
	\end{blockarray} 
	\ \to \ 
	\begin{blockarray}{c}
	\phantom{...}\\
	\begin{block}{(c)}
  	0   \\
  	1  \\
  	2  \\
  	0   \\
	\end{block}
	\end{blockarray}
 	\end{align*}
	\end{minipage}\quad
	\begin{minipage}{0.45\textwidth}
	\shadowbox{
	\parbox{\textwidth}{
	\vspace*{12pt}
	\begin{itemize}
	\item We can treat $D_{i,G_i}$ and $D_{G_i}$ as values of single variables!
	\item $P(D_{i,G_i})$ and $P(D_{i,G_i})$ can be modeled using $P_{NML}^1$-code.
	\end{itemize}}}  
	\end{minipage}

@@

\vskip-1.0ex
\begin{columns}[T]
  \begin{column}{0.63\paperwidth}
\begin{block}{References} %{{{
%\scriptsize
%\tiny
%{
%\bibliographystyle{apalike}
%\bibliography{../cosco}
%}

\begin{tiny}
\begin{thebibliography}{}
\bibitem[Liu et~al., 2012]{Liu2012}
Liu, Z. et~al. (2012).
\newblock Empirical evaluation of scoring functions for {B}ayesian network
  model selection.
\newblock {\em BMC Bioinformatics}, 13(15):S14.

\bibitem[Silander, T. et~al., 2007]{cosco.uai07}
Silander et~al. (2007).
\newblock On sensitivity of the {MAP} {Bayesian} network structure to the
  equivalent sample size parameter.
\newblock In {\em Proceedings of {UAI--07}}.

\bibitem[Silander, T. et~al., 2008]{cosco.pgm08a}
Silander et~al. (2008).
\newblock Factorized normalized maximum likelihood criterion for learning
  {B}ayesian network structures.
\newblock In {\em Proceedings of {PGM-08}}.

\bibitem[Suzuki, 2017]{Suzuki2017}
Suzuki, J. (2017).
\newblock A theoretical analysis of the {BDeu} scores in {B}ayesian network
  structure learning.
\newblock {\em Behaviormetrika}, 44(1):97--116.

\bibitem[Szpankowski and Weinberger, 2012]{Szpankowski2012}
Szpankowski, W. and Weinberger, M.~J. (2012).
\newblock Minimax pointwise redundancy for memoryless models over large
  alphabets.
\newblock {\em IEEE Transactions on Information Theory}, 58(7):4094--4104.
\end{thebibliography}
\end{tiny}

\end{block} %}}}
  \end{column}
  \begin{column}{0.33\paperwidth}
\begin{block}{Acknowledgements} %{{{
%\scriptsize
\tiny
{
J.L., E.J. and T.R. were supported by the Academy of Finland (COIN CoE and Project \textsc{TensorML}). J.L. was supported by the DoCS doctoral programme of the University of Helsinki.
}
\end{block} %}}}
  \end{column}
\end{columns}
\end{frame}
\end{document}
